### 2023-05-29 Manuscript1.Rmd
cleaning

---
title: "Manuscript"
author: "Jesper Fischer Ehmsen"
date: "`r Sys.Date()`"
output: html_document
---

### loading packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse, osfr, lmerTest, MuMIn, DHARMa, here, cmdstanr, tidybayes, bayesplot, posterior,future,furrr, patchwork, R.utils, brms, extraDistr,cowplot,HDInterval)

scripts = c("agent.R","parameter_recovery.R","stan_functions.R","utility functions.R", "model recovery.R","visualize_parameters.R")
q = lapply(here("scripts", scripts), source)
```

### parameter visualization for the four models (figure 1 & 2)
```{r parameter visualization, fig.height=10, fig.width=12}
dd = get_experiment()

rwplot = RW(dd)

kalmanplot = KALMAN_v2(dd)

wbplot = WB(dd)

RW_logistic = RW_logistic_vis(dd)

layout = (c(
  area(1, 1, 1,1),
  area(1, 2, 1,2)

))
plot(layout)

p1 = RW_logistic[[1]]+rwplot[[1]]+scale_y_continuous(" ")+plot_layout(design = layout)+plot_annotation(tag_levels = list(c('B', 'C')))
p2 = wbplot[[1]]+kalmanplot[[1]]+scale_y_continuous(" ")+plot_layout(design = layout)+plot_annotation(tag_levels = list(c("D","E")))
p3 = RW_logistic[[2]]+rwplot[[2]]+scale_y_continuous(" ")+plot_layout(design = layout)+plot_annotation(tag_levels = list(c("A","B")))
p4  = wbplot[[2]]+kalmanplot[[2]]+scale_y_continuous(" ")+plot_layout(design = layout)+plot_annotation(tag_levels = list(c("C","D")))

p1
p2
p3
p4

```

### parameter recovery for the four models
```{r parameter recovery for RW}
mu_alpha = seq(0.1,0.5,by = 0.1)
kappa_alpha = seq(10,300,length.out = 3)

mu_w1 = seq(0.1,0.5,by = 0.1)
kappa_w1 = seq(10,300,length.out = 3)

sd_precision_percept = seq(0.1,2, length.out = 3)

sd_beta = seq(0.1,2, length.out = 3)

nsubs = 10

parameters = expand.grid(mu_alpha = mu_alpha,
                         kappa_alpha = kappa_alpha,
                         mu_w1 = mu_w1,
                         kappa_w1 = kappa_w1,
                         sd_precision_percept = sd_precision_percept,
                         sd_beta = sd_beta,
                         nsubs = nsubs)

parameters$id = 1:nrow(parameters)

data_list <- split(parameters, parameters$id)

plan(multisession, workers = 3)

data_list1 = data_list[501:600]

rw_pp = furrr::future_map(data_list, ~parameter_recovery_rw(.x), .progress = TRUE, .options = furrr_options(seed = 123))
```

```{r parameter recovery for WB}

mu_alpha = seq(0.1,0.5,by = 0.1)
kappa_alpha = seq(10,300, length.out = 3)

mu_w1 = seq(0.1,0.5,by = 0.1)
kappa_w1 = seq(10,300, length.out = 3)

mu_w2 = seq(0.3,0.3,by = 0.1)
kappa_w2 = seq(10,300, length.out = 3)

sd_percept_precision = seq(0.1,2, length.out = 3)

sd_beta = seq(0.1,2, length.out = 3)

nsubs = 10

parameters = expand.grid(nsubs = nsubs, 
                         mu_alpha = mu_alpha, 
                         kappa_alpha = kappa_alpha, 
                         mu_w1 =  mu_w1, kappa_w1 = kappa_w1,
                         mu_w2 = mu_w2, kappa_w2 = kappa_w2, 
                         sd_percept_precision = sd_percept_precision, 
                         sd_beta = sd_beta)

parameters$id = 1:nrow(parameters)

data_list <- split(parameters, parameters$id)

plan(multisession, workers = availableCores()-1)

data_list = data_list[1:100]

wb_pp = furrr::future_map(data_list, ~parameter_recovery_wb(.x), .progress = TRUE, .options = furrr_options(seed = 123))

```

```{r parameter recovery for Kalman}

mu_sigmaEta = seq(0.1,1,length.out = 3)
mu_sigmaPsi = seq(0.1,5,length.out = 3)
mu_sigmaEpsilon = seq(0.1,5,length.out = 3)

sd_sigmaEta = seq(0.1,1,length.out = 2)
sd_sigmaPsi = seq(0.1,1,length.out = 2)
sd_sigmaEpsilon = seq(0.1,1,length.out = 2)

sd_precision_percept = seq(0.1,1, length.out = 2)

sd_beta = seq(0.1,1, length.out = 2)

nsubs = 10

parameters = expand.grid(mu_sigmaEta = mu_sigmaEta,
                         mu_sigmaPsi = mu_sigmaPsi,
                         mu_sigmaEpsilon = mu_sigmaEpsilon,
                         sd_sigmaEta = sd_sigmaEta,
                         sd_sigmaPsi = sd_sigmaPsi,
                         sd_sigmaEpsilon = sd_sigmaEpsilon,
                         sd_precision_percept = sd_precision_percept, 
                         sd_beta = sd_beta,
                         nsubs = nsubs)

parameters$id = 1:nrow(parameters)

data_list <- split(parameters, parameters$id)

#plan(multisession, workers = availableCores()-1)

data_list = data_list[1:2]

kalman_pp = furrr::future_map(data_list, ~parameter_recovery_kalman(.x), .progress = TRUE, .options = furrr_options(seed = 123))

```

```{r parameter recovery for logistic}
mu_alpha = seq(0.1,0.5, length.out = 5)
kappa_alpha = seq(10,100, length.out = 2)

mu_b0 = -1.5
sd_b0 = 0.5

mu_b1 = seq(0.1,2,length.out = 5)
sd_b1 = seq(0.1,2,length.out = 3)

mu_b2 = seq(0.1,2,length.out = 5)
sd_b2 = seq(0.1,2,length.out = 3)

mu_b0_b = -4
sd_b0_b = 0.5

mu_b1_b = 7
sd_b1_b = 0.5

mu_b2_b = 0.7
sd_b2_b = 0.5

sd_percept_precision = 0.8

nsubs = 10

parameters = expand.grid(mu_alpha = mu_alpha,
                         kappa_alpha = kappa_alpha,
                         mu_b0 = mu_b0,
                         sd_b0 = sd_b0,
                         mu_b1 = mu_b1,
                         sd_b1 = sd_b1,
                         mu_b2 = mu_b2,
                         sd_b2 = sd_b2,
                         mu_b0_b = mu_b0_b,
                         sd_b0_b = sd_b0_b,
                         mu_b1_b = mu_b1_b,
                         sd_b1_b = sd_b1_b,
                         mu_b2_b = mu_b2_b,
                         sd_b2_b = sd_b2_b,
                         sd_percept_precision = sd_percept_precision,
                         nsubs = nsubs)

parameters$id = 1:nrow(parameters)

data_list <- split(parameters, parameters$id)

plan(multisession, workers = 5)

rw_pp = furrr::future_map(data_list, ~parameter_recovery_logistic_rw(.x), .progress = TRUE, .options = furrr_options(seed = 123))

```

#plotting parameter recovery (figure 3)
```{r}
files = paste0(here::here("workspace","parameterrecovery"),"/",list.files(here::here("workspace","parameterrecovery")))

#plot parameterrecovery
load(files[[1]])
q = lapply(here("scripts", scripts), source)
kalmanplot = plot_parameterrecovery(kalman_pp, div = T, subs = 5)
load(files[[2]])
q = lapply(here("scripts", scripts), source)
logisticplot = plot_parameterrecovery(rw_pp_logs, div = F, subs = 5)
load(files[[3]])
q = lapply(here("scripts", scripts), source)
rwplot = plot_parameterrecovery(rw_pp, div = T, subs = 5)
load(files[[4]])
q = lapply(here("scripts", scripts), source)
wbplot = plot_parameterrecovery(wb_pp, div = T, subs = 5)

layout = (c(
  area(1, 1, 2,3),
  area(3, 1,5,3),
  area(6, 1,6,3)

))

((rwplot[[1]]+wbplot[[1]])/(kalmanplot[[1]]+logisticplot[[1]]))+
  plot_annotation(tag_levels = 'A')+
  plot_layout(guides = 'collect')

((rwplot[[2]]+wbplot[[2]])/(kalmanplot[[2]]+logisticplot[[2]]))+
  plot_annotation(tag_levels = 'A')+
  plot_layout(guides = 'collect')

```

### model recovery for the three computation models
```{r model recovery}
########################################### Resorla wagner parameters:
mu_alpha = rep(0.3,100)
kappa_alpha = 50
mu_w1 = 0.3
kappa_w1 = 50
sd_precision_percept = 0.5
sd_beta = 0.5
nsubs = 10
parameters = expand.grid(mu_alpha = mu_alpha,
                         kappa_alpha = kappa_alpha,
                         mu_w1 = mu_w1,
                         kappa_w1 = kappa_w1,
                         sd_precision_percept = sd_precision_percept,
                         sd_beta = sd_beta,
                         nsubs = nsubs)

parameters$id = 1:nrow(parameters)
data_list_rw <- split(parameters, parameters$id)

######################################## Weighted bayes

mu_alpha = rep(0.3,100)
kappa_alpha = 50
mu_w1 = 0.3
kappa_w1 = 50
mu_w2 = 0.3
kappa_w2 = 50
sd_percept_precision = 0.5
sd_beta = 0.5
nsubs = 10
parameters = expand.grid(nsubs = nsubs, 
                         mu_alpha = mu_alpha, 
                         kappa_alpha = kappa_alpha, 
                         mu_w1 =  mu_w1, kappa_w1 = kappa_w1,
                         mu_w2 = mu_w2, kappa_w2 = kappa_w2, 
                         sd_percept_precision = sd_percept_precision, 
                         sd_beta = sd_beta)
parameters$id = 1:nrow(parameters)
data_list_wb <- split(parameters, parameters$id)

######################################### logistic RW

mu_alpha = rep(0.3,100)
kappa_alpha = 50

mu_b0 = -1.5
sd_b0 = 0.5

mu_b1 = 2
sd_b1 = 1

mu_b2 = 1
sd_b2 = 1

mu_b0_b = -3
sd_b0_b = 0.5

mu_b1_b = 5
sd_b1_b = 0.5

mu_b2_b = 0.7
sd_b2_b = 0.5

sd_percept_precision = 0.5

nsubs = 10

parameters = expand.grid(mu_alpha = mu_alpha,
                         kappa_alpha = kappa_alpha,
                         mu_b0 = mu_b0,
                         sd_b0 = sd_b0,
                         mu_b1 = mu_b1,
                         sd_b1 = sd_b1,
                         mu_b2 = mu_b2,
                         sd_b2 = sd_b2,
                         mu_b0_b = mu_b0_b,
                         sd_b0_b = sd_b0_b,
                         mu_b1_b = mu_b1_b,
                         sd_b1_b = sd_b1_b,
                         mu_b2_b = mu_b2_b,
                         sd_b2_b = sd_b2_b,
                         sd_percept_precision = sd_percept_precision,
                         nsubs = nsubs)

parameters$id = 1:nrow(parameters)

data_list_logs <- split(parameters, parameters$id)

######################################### fit data_rw:
#loos = model_fitter_rwdata(data_list_rw[[1]])

data_list = data_list_rw

plan(multisession, workers = 10)

rw_loo = furrr::future_map(data_list, ~model_fitter_rwdata(.x), .progress = TRUE, .options = furrr_options(seed = 123))

elpld_loo <- map_dfr(rw_loo, ~ .x[[1]])

full_loo <- map_dfr(rw_loo, ~ .x[[2]])

########################################## fit data_wb

data_list = data_list_wb

plan(multisession, workers = availableCores()-1)

wb_loo = furrr::future_map(data_list, ~model_fitter_wbdata(.x), .progress = TRUE, .options = furrr_options(seed = 123))

elpld_loo <- map_dfr(wb_loo, ~ .x[[1]])

full_loo <- map_dfr(wb_loo, ~ .x[[2]])

########################################### fit kalman data

data_list = data_list_logs[1:50]

plan(multisession, workers = 10)

kalman_loo = furrr::future_map(data_list, ~model_fitter_kalmandata(.x), .progress = TRUE, .options = furrr_options(seed = 123))

elpld_loo <- map_dfr(kalman_loo, ~ .x[[1]])

full_loo <- map_dfr(kalman_loo, ~ .x[[2]])

```

### table 1
```{r model recovery table}

files = paste0(here::here("workspace","model recovery"),"/",list.files(here::here("workspace","model recovery")))

#plot parameterrecovery
load(files[[1]])

get_value = function(loo){
  get_from_means <- do.call(rbind, lapply(loo, "[[", 1))

  get_from_means$id = as.factor(get_from_means$id)
# Group by ID and find the model with the minimum value
  min_values <- get_from_means %>%
  group_by(id) %>%
  filter(means == min(means))

# Count the occurrences of each model with the minimum value per run
  model_counts <- min_values %>%
    group_by(names) %>%
    summarize(Count = n())

  get_from_loo <- do.call(rbind, lapply(loo, "[[", 2))

  get_from_loo = get_from_loo %>% filter(elpd_diff == 0) %>% mutate(model = row.names(.)) %>% mutate(realmodel = sub("^[0-9]+\\.(.*)", "\\1", model)) %>% group_by(realmodel) %>% summarize(Count = n())

  return(list(fromloo = get_from_loo, frommean = model_counts))
}

kalmandata = get_value(kalman_loo)
rwdata = get_value(rw_loo)
wbdata = get_value(wb_loo)

```

### fitting the models to session 1's data

```{r fitting RW model to session 1}
df= read.csv(here::here("data_TPL_AC.csv")) %>% 
  rename_all(tolower) %>% 
  filter(trial != 0) %>% 
  rename(id = participant, stim = stimulus, percept = rating, pred = prediction, percept_bin = quality) %>% 
  mutate_all(replace_na, 0)

df$u = ifelse(df$cue == 1 & df$stim == 1, 1, ifelse(df$cue == 0 & df$stim == 0, 1,0))

df = df %>% filter(session == 1)

#fitting hierachically
data = df %>% mutate(percept = percept/100)

data = data %>% mutate(across(percept, ~case_when(
    . < 0.001 ~ 0.001,
    . > 0.999 ~ 0.999,
    TRUE ~ .
  )))

data1 = list(nsubs = length(unique(data$id)),
            ntrials = nrow(data %>% filter(id %in% unique(id)[1])),
            percept = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept)%>% mutate(trial= NULL)),
            expectPain = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = pred)%>% mutate(trial= NULL)),
            percept_bin = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept_bin)%>% mutate(trial= NULL)),
            stim = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = stim)%>% mutate(trial= NULL)),
            cues = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = cue)%>% mutate(trial= NULL)))

mod = cmdstan_model(here::here("Stan","myRW_real.stan"))

rw_fit_sess1 <- mod$sample(
      data = data1,
      chains = 4, 
      parallel_chains = 4,
      refresh = 100,
      adapt_delta = 0.90,
      max_treedepth = 12
    )

ppu_hier(fit = rw_fit_sess1,
         parameters = c("alpha","w1","precision_percept"),
         reals = NA)

```

```{r fitting WB model to session 1}
df= read.csv(here::here("data_TPL_AC.csv")) %>% 
  rename_all(tolower) %>% 
  filter(trial != 0) %>% 
  rename(id = participant, stim = stimulus, percept = rating, pred = prediction, percept_bin = quality) %>% 
  mutate_all(replace_na, 0)

df$u = ifelse(df$cue == 1 & df$stim == 1, 1, ifelse(df$cue == 0 & df$stim == 0, 1,0))

df = df %>% filter(session == 1)

#fitting hierachically
data = df %>% mutate(percept = percept/100)

data = data %>% mutate(across(c(percept,stim), ~case_when(
    . < 0.001 ~ 0.001,
    . > 0.999 ~ 0.999,
    TRUE ~ .
  )))

data1 = list(nsubs = length(unique(data$id)),
            ntrials = nrow(data %>% filter(id %in% unique(id)[1])),
            percept = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept)%>% mutate(trial= NULL)),
            pred = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = pred)%>% mutate(trial= NULL)),
            percept_bin = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept_bin)%>% mutate(trial= NULL)),
            stim = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = stim)%>% mutate(trial= NULL)),
            cue = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = cue)%>% mutate(trial= NULL))
            )

mod = cmdstan_model(here::here("Stan","hier_weighted_bayes.stan"))

wb_fit_sess1  <- mod$sample(
      data = data1,
      chains = 4, 
      parallel_chains = 4,
      refresh = 100,
      adapt_delta = 0.90,
      max_treedepth = 12
    )

wb_fit_sess1$diagnostic_summary()

alpha = get_diag(wb_fit_sess1, "alpha")
w1 = get_diag(wb_fit_sess1, "w1")
w2 = get_diag(wb_fit_sess1, "w2")
pp = get_diag(wb_fit_sess1, "percept_precision")
beta = get_diag(wb_fit_sess1, "beta")

```

```{r, fitting to real data tester hierachical logs RW sessions1}

df= read.csv(here::here("data_TPL_AC.csv")) %>%
  rename_all(tolower) %>%
  filter(trial != 0) %>%
  rename(id = participant, stim = stimulus, percept = rating, pred = prediction, percept_bin = quality) %>%
  mutate_all(replace_na, 0)

df$u = ifelse(df$cue == 1 & df$stim == 1, 1, ifelse(df$cue == 0 & df$stim == 0, 1,0))
df$pred = ifelse(df$cue == 1 & df$pred == 1, 1, ifelse(df$cue == 0 & df$pred == 0, 1,0))

df = df %>% filter(session == 1)

#source(here::here("stan_functions.R"))
#fitting hierachically
data = df %>% mutate(percept = percept/100)

data = data %>% mutate(across(percept, ~case_when(
    . < 0.001 ~ 0.001,
    . > 0.999 ~ 0.999,
    TRUE ~ .
  )))

data2 = list(nsubs = length(unique(data$id)),
            ntrials = nrow(data %>% filter(id %in% unique(id)[1])),
            pred = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = pred)%>% mutate(trial= NULL)),
            percept_bin = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept_bin)%>% mutate(trial= NULL)),
            percept = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept)%>% mutate(trial= NULL)),
            stim = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = stim)%>% mutate(trial= NULL)),
            u = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = u)%>% mutate(trial= NULL)),
            cues = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = cue)%>% mutate(trial= NULL))
            )

mod = cmdstan_model(here::here("Stan","rw_vs_rw_hier_complete.stan"))

rw_logs_fit_ses1 <- mod$sample(
      data = data2,
      chains = 4,
      parallel_chains = 4,
      refresh = 100,
      adapt_delta = 0.90,
      max_treedepth = 12
    )

sum1 = rw_logs_fit_ses1$summary()
```

### Model comparison on sessions 1 (table 2)
```{r, model comparison}
#rw = rw_fit_sess1$loo()
#wb = wb_fit_sess1$loo()
#rw_logs = rw_logs_fit_ses1$loo()

rw = as_draws_df(rw_fit_sess1$draws(variables = "log_lik"))
wb = as_draws_df(wb_fit_sess1$draws(variables = "log_lik"))
rw_logs = as_draws_df(rw_logs_fit_ses1$draws(variables = "log_lik"))

rows_with_na <- which(apply(is.na(wb), 1, any))
wb = wb[-rows_with_na,]
rw = rw[-rows_with_na,]
rw_logs = rw_logs[-rows_with_na, ]

wb = loo::loo(as.matrix(wb))
rw = loo::loo(as.matrix(rw))
rw_logs = loo::loo(as.matrix(rw_logs))

loo_compare(rw,wb,rw_logs)
loo_model_weights(list(rw,wb,rw_logs))
```

### fitting the models to session 2's data

```{r fitting RW model to session 2}
df= read.csv(here::here("data_TPL_AC.csv")) %>% 
  rename_all(tolower) %>% 
  filter(trial != 0) %>% 
  rename(id = participant, stim = stimulus, percept = rating, pred = prediction, percept_bin = quality) %>% 
  mutate_all(replace_na, 0)

df$u = ifelse(df$cue == 1 & df$stim == 1, 1, ifelse(df$cue == 0 & df$stim == 0, 1,0))

df = df %>% filter(session == 2)

#fitting hierachically
data = df %>% mutate(percept = percept/100)

data = data %>% mutate(across(percept, ~case_when(
    . < 0.001 ~ 0.001,
    . > 0.999 ~ 0.999,
    TRUE ~ .
  )))

data1 = list(nsubs = length(unique(data$id)),
            ntrials = nrow(data %>% filter(id %in% unique(id)[1])),
            percept = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept)%>% mutate(trial= NULL)),
            expectPain = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = pred)%>% mutate(trial= NULL)),
            percept_bin = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept_bin)%>% mutate(trial= NULL)),
            stim = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = stim)%>% mutate(trial= NULL)),
            cues = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = cue)%>% mutate(trial= NULL)))

mod = cmdstan_model(here::here("Stan","myRW_real.stan"))

rw_fit_sess2 <- mod$sample(
      data = data1,
      chains = 4, 
      parallel_chains = 4,
      refresh = 100,
      adapt_delta = 0.90,
      max_treedepth = 12
    )

ppu_hier(fit = rw_fit_sess2,
         parameters = c("alpha","w1","precision_percept"),
         reals = NA)

```

### fitting the models to session 2's data

```{r fitting WB model to session 2}

df= read.csv(here::here("data_TPL_AC.csv")) %>% 
  rename_all(tolower) %>% 
  filter(trial != 0) %>% 
  rename(id = participant, stim = stimulus, percept = rating, pred = prediction, percept_bin = quality) %>% 
  mutate_all(replace_na, 0)

df$u = ifelse(df$cue == 1 & df$stim == 1, 1, ifelse(df$cue == 0 & df$stim == 0, 1,0))

df = df %>% filter(session == 2)

#fitting hierachically
data = df %>% mutate(percept = percept/100)

data = data %>% mutate(across(c(percept,stim), ~case_when(
    . < 0.001 ~ 0.001,
    . > 0.999 ~ 0.999,
    TRUE ~ .
  )))

data1 = list(nsubs = length(unique(data$id)),
            ntrials = nrow(data %>% filter(id %in% unique(id)[1])),
            percept = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept)%>% mutate(trial= NULL)),
            pred = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = pred)%>% mutate(trial= NULL)),
            percept_bin = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept_bin)%>% mutate(trial= NULL)),
            stim = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = stim)%>% mutate(trial= NULL)),
            cue = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = cue)%>% mutate(trial= NULL))
            )

mod = cmdstan_model(here::here("Stan","hier_weighted_bayes.stan"))

wb_fit_sess2 <- mod$sample(
      data = data1,
      chains = 4, 
      parallel_chains = 4,
      refresh = 100,
      adapt_delta = 0.90,
      max_treedepth = 12
    )

wb_fit_sess2$diagnostic_summary()

alpha = get_diag(wb_fit_sess1, "alpha")
w1 = get_diag(wb_fit_sess1, "w1")
w2 = get_diag(wb_fit_sess1, "w2")
pp = get_diag(wb_fit_sess1, "percept_precision")
beta = get_diag(wb_fit_sess1, "beta")

```

### fitting the models to session 2's data

```{r, fitting to real data tester hierachical logs RW sessions2}

df= read.csv(here::here("data_TPL_AC.csv")) %>%
  rename_all(tolower) %>%
  filter(trial != 0) %>%
  rename(id = participant, stim = stimulus, percept = rating, pred = prediction, percept_bin = quality) %>%
  mutate_all(replace_na, 0)

df$u = ifelse(df$cue == 1 & df$stim == 1, 1, ifelse(df$cue == 0 & df$stim == 0, 1,0))
df$pred = ifelse(df$cue == 1 & df$pred == 1, 1, ifelse(df$cue == 0 & df$pred == 0, 1,0))

df = df %>% filter(session == 2)

#source(here::here("stan_functions.R"))
#fitting hierachically
data = df %>% mutate(percept = percept/100)

data = data %>% mutate(across(percept, ~case_when(
    . < 0.001 ~ 0.001,
    . > 0.999 ~ 0.999,
    TRUE ~ .
  )))

data2 = list(nsubs = length(unique(data$id)),
            ntrials = nrow(data %>% filter(id %in% unique(id)[1])),
            pred = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = pred)%>% mutate(trial= NULL)),
            percept_bin = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept_bin)%>% mutate(trial= NULL)),
            percept = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept)%>% mutate(trial= NULL)),
            stim = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = stim)%>% mutate(trial= NULL)),
            u = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = u)%>% mutate(trial= NULL)),
            cues = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = cue)%>% mutate(trial= NULL))
            )

mod = cmdstan_model(here::here("Stan","rw_vs_rw_hier_complete.stan"))

rw_logs_fit_ses2 <- mod$sample(
      data = data2,
      chains = 4,
      parallel_chains = 4,
      refresh = 100,
      adapt_delta = 0.90,
      max_treedepth = 12
    )

```

### Model comparison on sessions 2 (table 2)

```{r, model comparison}
#rw = rw_fit_sess1$loo()
#wb = wb_fit_sess1$loo()
#rw_logs = rw_logs_fit_ses1$loo()

rw = as_draws_df(rw_fit_sess2$draws(variables = "log_lik"))
wb = as_draws_df(wb_fit_sess2$draws(variables = "log_lik"))
rw_logs = as_draws_df(rw_logs_fit_ses2$draws(variables = "log_lik"))

rows_with_na <- which(apply(is.na(wb), 1, any))
wb = wb[-rows_with_na,]
rw = rw[-rows_with_na,]
rw_logs = rw_logs[-rows_with_na, ]

wb = loo::loo(as.matrix(wb))
rw = loo::loo(as.matrix(rw))
rw_logs = loo::loo(as.matrix(rw_logs))

loo_compare(rw_logs,rw,wb)

loo_model_weights(list(rw_logs,rw,wb))
```

### Fitting the models to both sessions

```{r hierachical session correlation (RW)}
df= read.csv(here::here("data_TPL_AC.csv")) %>%
  rename_all(tolower) %>%
  filter(trial != 0) %>%
  rename(id = participant, stim = stimulus, percept = rating, pred = prediction, percept_bin = quality) %>%
  mutate_all(replace_na, 0) %>% mutate(id = as.factor(id), trial = as.factor(trial), session = as.factor(session))

df$u = ifelse(df$cue == 1 & df$stim == 1, 1, ifelse(df$cue == 0 & df$stim == 0, 1,0))

#fitting hierachically
data = df %>% mutate(percept = percept/100)

data = data %>% mutate(across(percept, ~case_when(
    . < 0.001 ~ 0.001,
    . > 0.999 ~ 0.999,
    TRUE ~ .
  )))

data1 = list(nsubs = length(unique(data$id)),
            ntrials = nrow(data %>% filter(session == 1) %>% filter(id %in% unique(id)[1])),
            sesions = length(unique(data$session)),
            percept = get_3darray(data,variable = "percept"),
            expectPain = get_3darray(data,variable = "pred"),
            percept_bin = get_3darray(data,variable = "percept_bin"),
            stim = get_3darray(data,variable = "stim"),
            cues = get_3darray(data,variable = "cue"),
            u = get_3darray(data,variable = "u")
            )

mod = cmdstan_model(here::here("Stan","fitting both sessions","myRW_real_both_sess.stan"))

fit_rw_bs <- mod$sample(
  data = data1,
  chains = 4,
  parallel_chains = 4,
  refresh = 100,
  adapt_delta = 0.90,
  max_treedepth = 12
)

alpha_wb = get_corplotreal(fit_rw_bs, "alpha", ymin = 0.1, ymax = 0.2, xmin = 0.2, xmax = 0.3)
w1_wb = get_corplotreal(fit_rw_bs, "w1", ymin = 0.125, ymax = 0.3, xmin = 0.6, xmax = 0.75)

HDInterval::hdi(as_draws_df(fit_rw_bs$draws(variables = "correlation_alpha")) %>% select(starts_with("correlation")))

HDInterval::hdi(as_draws_df(fit_rw_bs$draws(variables = "correlation_w1")) %>% select(starts_with("correlation")))

```

```{r hierachical session correlation (WB)}

df= read.csv(here::here("data_TPL_AC.csv")) %>%
  rename_all(tolower) %>%
  filter(trial != 0) %>%
  rename(id = participant, stim = stimulus, percept = rating, pred = prediction, percept_bin = quality) %>%
  mutate_all(replace_na, 0) %>% mutate(id = as.factor(id), trial = as.factor(trial), session = as.factor(session))

df$u = ifelse(df$cue == 1 & df$stim == 1, 1, ifelse(df$cue == 0 & df$stim == 0, 1,0))

#fitting hierachically
data = df %>% mutate(percept = percept/100)

data = data %>% mutate(across(c(percept,stim), ~case_when(
    . < 0.001 ~ 0.001,
    . > 0.999 ~ 0.999,
    TRUE ~ .
  )))

data1 = list(nsubs = length(unique(data$id)),
            ntrials = nrow(data %>% filter(session == 1) %>% filter(id %in% unique(id)[1])),
            sesions = length(unique(data$session)),
            percept = get_3darray(data,variable = "percept"),
            pred = get_3darray(data,variable = "pred"),
            percept_bin = get_3darray(data,variable = "percept_bin"),
            stim = get_3darray(data,variable = "stim"),
            cue = get_3darray(data,variable = "cue")
            )

mod = cmdstan_model(here::here("Stan","fitting both sessions","hier_weighted_bayes_correlation.stan"))

fit_wb_bs <- mod$sample(
  data = data1,
  chains = 4,
  parallel_chains = 4,
  refresh = 50,
  adapt_delta = 0.90,
  max_treedepth = 12
)

fit_wb_bs$diagnostic_summary()

correlation = as_draws_df(fit_wb_bs$draws(variables = "correlation_alpha"))
hist(correlation$correlation_alpha)
correlation = as_draws_df(fit_wb_bs$draws(variables = "correlation_w1"))
hist(correlation$correlation_w1)
correlation = as_draws_df(fit_wb_bs$draws(variables = "correlation_w2"))
hist(correlation$correlation_w2)

alphaplot1 = get_corplot(wb_fit_sess1, wb_fit_sess2, "alpha")

alpha_wb = get_corplotreal(fit_wb_bs, "alpha", ymin = -0.075, ymax = 0.325, xmin = 0.65, xmax = 1)
w1_wb = get_corplotreal(fit_wb_bs, "w1", ymin = 0.02, ymax = 0.12, xmin = 0.19, xmax = 0.27)
w2_wb = get_corplotreal(fit_wb_bs, "w2", ymin = 0.1, ymax = 0.2, xmin = 0.4, xmax = 0.6)

HDInterval::hdi(as_draws_df(fit_wb_bs$draws(variables = "correlation_alpha")) %>% select(starts_with("correlation")))

HDInterval::hdi(as_draws_df(fit_wb_bs$draws(variables = "correlation_w1")) %>% select(starts_with("correlation")))

HDInterval::hdi(as_draws_df(fit_wb_bs$draws(variables = "correlation_w2")) %>% select(starts_with("correlation")))

```

```{r logistic both sessions}
df= read.csv(here::here("data_TPL_AC.csv")) %>%
  rename_all(tolower) %>%
  filter(trial != 0) %>%
  rename(id = participant, stim = stimulus, percept = rating, pred = prediction, percept_bin = quality) %>%
  mutate_all(replace_na, 0) %>% mutate(id = as.factor(id), trial = as.factor(trial), session = as.factor(session))

df$u = ifelse(df$cue == 1 & df$stim == 1, 1, ifelse(df$cue == 0 & df$stim == 0, 1,0))
df$pred = ifelse(df$cue == 1 & df$pred == 1, 1, ifelse(df$cue == 0 & df$pred == 0, 1,0))

#fitting hierachically
data = df %>% mutate(percept = percept/100)

data = data %>% mutate(across(percept, ~case_when(
    . < 0.001 ~ 0.001,
    . > 0.999 ~ 0.999,
    TRUE ~ .
  )))

data1 = list(nsubs = length(unique(data$id)),
            ntrials = nrow(data %>% filter(session == 1) %>% filter(id %in% unique(id)[1])),
            sesions = length(unique(data$session)),
            percept = get_3darray(data,variable = "percept"),
            pred = get_3darray(data,variable = "pred"),
            percept_bin = get_3darray(data,variable = "percept_bin"),
            stim = get_3darray(data,variable = "stim"),
            cues = get_3darray(data,variable = "cue"),
            u = get_3darray(data,variable = "u")
            )

mod = cmdstan_model(here::here("Stan","fitting both sessions","rw_vs_rw_hier_complete_both_sessions.stan"))

fit_rw_logs_bs <- mod$sample(
  data = data1,
  chains = 4,
  parallel_chains = 4,
  refresh = 50,
  adapt_delta = 0.90,
  max_treedepth = 12
)

HDInterval::hdi(get_correlation(rw_logs_fit_ses1,rw_logs_fit_ses2,"alpha"))

HDInterval::hdi(get_correlation(rw_logs_fit_ses1,rw_logs_fit_ses2,"b1"))

HDInterval::hdi(get_correlation(rw_logs_fit_ses1,rw_logs_fit_ses2,"b2"))

```

### plotting the correlation (figure 4)

```{r correlation plots}

#logs

alpha_logs = get_corplot(rw_logs_fit_ses1,rw_logs_fit_ses2,"alpha",ymin = 0, ymax = 0.3, xmin = 0.55, xmax = 0.85)
#b0_logs = get_corplot(rw_logs_fit_ses1,rw_logs_fit_ses2,"b0",ymin = -4.25, ymax = -3, xmin = -3, xmax = -2)
b1_logs = get_corplot(rw_logs_fit_ses1,rw_logs_fit_ses2,"b1",ymin = 0.2, ymax = 1.5, xmin = 2.5, xmax = 3.4)
b2_logs = get_corplot(rw_logs_fit_ses1,rw_logs_fit_ses2,"b2",ymin = -0.2, ymax = 0, xmin = 0.3, xmax = 0.575)

rw_logs = alpha_logs+b1_logs+b2_logs
#wb
alpha_wb = get_corplotreal(fit_wb_bs, "alpha", ymin = -0.075, ymax = 0.275, xmin = 0.56, xmax = 0.91)
w1_wb = get_corplotreal(fit_wb_bs, "w1", ymin = 0.02, ymax = 0.1, xmin = 0.17, xmax = 0.25)
w2_wb = get_corplotreal(fit_wb_bs, "w2", ymin = 0.1, ymax = 0.21, xmin = 0.3, xmax = 0.55)

wb = alpha_wb+w1_wb+w2_wb
#rw
alpha_rw = get_corplotreal(fit_rw_bs, "alpha", ymin = 0.1, ymax = 0.2, xmin = 0.2, xmax = 0.3)
w1_rw = get_corplotreal(fit_rw_bs, "w1", ymin = 0.125, ymax = 0.3, xmin = 0.6, xmax = 0.75)

rw = alpha_rw+w1_rw

rw+plot_annotation(tag_levels = list(c("A",'B')))
wb+plot_annotation(tag_levels = list(c("C",'D',"E")))
rw_logs+plot_annotation(tag_levels = list(c("F",'G',"H")))
```

### fitting the kalman model just to see

```{r fitting Kalman model to session 1}
df= read.csv(here::here("data_TPL_AC.csv")) %>%
  rename_all(tolower) %>%
  filter(trial != 0) %>%
  rename(id = participant, stim = stimulus, percept = rating, pred = prediction, percept_bin = quality) %>%
  mutate_all(replace_na, 0)

df$u = ifelse(df$cue == 1 & df$stim == 1, 1, ifelse(df$cue == 0 & df$stim == 0, 1,0))

df = df %>% filter(session == 1)

#fitting hierachically
data = df %>% mutate(percept = percept/100)

data = data %>% mutate(across(percept, ~case_when(
    . < 0.001 ~ 0.001,
    . > 0.999 ~ 0.999,
    TRUE ~ .
  )))

data1 = list(nsubs = length(unique(data$id)),
            ntrials = nrow(data %>% filter(id %in% unique(id)[1])),
            percept = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept)%>% mutate(trial= NULL)),
            expectPain = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = pred)%>% mutate(trial= NULL)),
            percept_bin = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept_bin)%>% mutate(trial= NULL)),
            stim = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = stim)%>% mutate(trial= NULL)),
            cues = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = cue)%>% mutate(trial= NULL)),
            u = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = u)%>% mutate(trial= NULL))
            )

mod = cmdstan_model(here::here("Stan","myKalmanfilter_v3_nocent.stan"))

kalman_fit_sess1 <- mod$sample(
      data = data1,
      chains = 4,
      parallel_chains = 4,
      refresh = 100,
      adapt_delta = 0.90,
      max_treedepth = 12
    )

ppu_hier(fit = kalman_fit_sess1,
         parameters = c("sigmaEta","sigmaPsi","sigmaEpsilon","precision_percept"),
         reals = NA)

# 
# ppu_sub(fit = fit,
#         parameter = "precision_percept",
#         reals = df,
#         lims = -1,
#         lime = 10,
#         nsubs = NA
#         )

get_diag(kalman_fit_sess1, "sigmaEta")
get_diag(kalman_fit_sess1, "sigmaPsi")
get_diag(kalman_fit_sess1, "sigmaEpsilon")

sum = kalman_fit_sess1$summary()

ass = sum %>% filter(str_detect(variable, "perceptmuu"))
ass = ass[(161*2-1):(161*3-1),]

exp = sum %>% filter(str_detect(variable, "exp_muu"))
exp = exp[(161*2-1):(161*3-1),]

ass$exp = exp$mean

ass %>% ggplot(aes(x=exp, y = mean))+geom_point()

```

```{r fitting Kalman model to session 2}
df= read.csv(here::here("data_TPL_AC.csv")) %>%
  rename_all(tolower) %>%
  filter(trial != 0) %>%
  rename(id = participant, stim = stimulus, percept = rating, pred = prediction, percept_bin = quality) %>%
  mutate_all(replace_na, 0)

df$u = ifelse(df$cue == 1 & df$stim == 1, 1, ifelse(df$cue == 0 & df$stim == 0, 1,0))

df = df %>% filter(session == 2)

#fitting hierachically
data = df %>% mutate(percept = percept/100)

data = data %>% mutate(across(percept, ~case_when(
    . < 0.001 ~ 0.001,
    . > 0.999 ~ 0.999,
    TRUE ~ .
  )))

data1 = list(nsubs = length(unique(data$id)),
            ntrials = nrow(data %>% filter(id %in% unique(id)[1])),
            percept = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept)%>% mutate(trial= NULL)),
            expectPain = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = pred)%>% mutate(trial= NULL)),
            percept_bin = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = percept_bin)%>% mutate(trial= NULL)),
            stim = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = stim)%>% mutate(trial= NULL)),
            cues = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = cue)%>% mutate(trial= NULL)),
            u = as.matrix(data %>% pivot_wider(id_cols = trial, names_from = id, values_from = u)%>% mutate(trial= NULL))
            )

mod = cmdstan_model(here::here("Stan","myKalmanfilter_v3_nocent.stan"))

kalman_fit_sess2 <- mod$sample(
      data = data1,
      chains = 4,
      parallel_chains = 4,
      refresh = 100,
      adapt_delta = 0.90,
      max_treedepth = 12
    )

ppu_hier(fit = kalman_fit_sess2,
         parameters = c("sigmaEta","sigmaPsi","sigmaEpsilon","precision_percept"),
         reals = NA)

```